---
title: "STAT 294 - Contemporary Topics in Statistics"
Source: World Bank https://data.worldbank.org"
author: "Martina Le-Bert"
output: flexdashboard::flex_dashboard
---


```{r,warning=FALSE, message=FALSE, echo=FALSE}
library(rnaturalearth) #to hold and facilitate interaction with Natural Earth map data.
library(wbstats) # to search and download data from the World Bank API.
library(leaflet) # to create and customize interactive maps.
library(DT) # provides an R interface to the JavaScript library DataTables.
library(ggplot2) # to data visualization.
library(dplyr) #to data manipulation.
library(wesanderson) #color palette

map <- ne_countries()
names(map)[names(map) == "iso_a3"] <- "ISO3"
names(map)[names(map) == "name"] <- "NAME"
```

```{r,warning=FALSE, message=FALSE, echo=FALSE}
e <- wb(
  indicator = "EN.ATM.GHGT.KT.CE",
  startdate = 2017, enddate = 2017
)

f <- wb(
  indicator = "NY.GDP.PCAP.PP.CD",
  startdate = 2017, enddate = 2017
)

g <- wb(
  indicator = "SP.DYN.LE00.IN",
startdate = 2017, enddate = 2017
)
```


```{r}
map$GHG <- e[match(map$ISO3, e$iso3), "value"]
map$GDP <- f[match(map$ISO3, f$iso3), "value"]
map$Life <- g[match(map$ISO3,g$iso3), "value"]
```



  
Row {.tabset .tabset-fade}
-------------------------------------

### Annual GHG emissions, 2017



```{r}
pal <- colorBin(
  palette = c("#DD8D29", "#E2D200", "#46ACC8", "#E58601", "#B40F20"), domain = map$GHG,
  bins = seq(800, max(map$GHG, na.rm = TRUE) + 10, length.out=10)
)

map$labels <- paste0(
  "<strong> Country: </strong> ",
  map$NAME, "<br/> ",
  "<strong> GHG: </strong> ",
  map$GHG, "<br/> "
) %>%
  lapply(htmltools::HTML)

leaflet(map) %>%
  addTiles() %>%
  setView(lng = 0, lat = 30, zoom = 2) %>%
  addPolygons(
    fillColor = ~ pal(GHG),
    color = "white",
    fillOpacity = 0.7,
    label = ~labels,
    highlight = highlightOptions(
      color = "black",
      bringToFront = TRUE
    )
  ) %>%
  leaflet::addLegend(
    pal = pal, values = ~GHG,
    opacity = 0.7, title = "GHG"
  )
```

### GDP per capita, 2017

```{r}
pal <- colorBin(
  palette = c("#DD8D29", "#E2D200", "#46ACC8", "#E58601", "#B40F20"), domain = map$GDP,
  bins = seq(770, max(map$GDP, na.rm = TRUE) + 10, length.out=10)
)

map$labels <- paste0(
  "<strong> Country: </strong> ",
  map$NAME, "<br/> ",
  "<strong> GDP: </strong> ",
  map$GDP, "<br/> "
) %>%
  lapply(htmltools::HTML)

leaflet(map) %>%
  addTiles() %>%
  setView(lng = 0, lat = 30, zoom = 2) %>%
  addPolygons(
    fillColor = ~ pal(GDP),
    color = "white",
    fillOpacity = 0.7,
    label = ~labels,
    highlight = highlightOptions(
      color = "black",
      bringToFront = TRUE
    )
  ) %>%
  leaflet::addLegend(
    pal = pal, values = ~GDP,
    opacity = 0.7, title = "GDP"
  )
```

### Life expectancy, 2017

```{r}
pal <- colorBin(
  palette = c("#DD8D29", "#E2D200", "#46ACC8", "#E58601", "#B40F20"), domain = map$Life,
  bins = seq(50, max(map$Life, na.rm = TRUE) + 10, length.out=10)
)

map$labels <- paste0(
  "<strong> Country: </strong> ",
  map$NAME, "<br/> ",
  "<strong> Life Expectancy: </strong> ",
  map$Life, "<br/> "
) %>%
  lapply(htmltools::HTML)

leaflet(map) %>%
  addTiles() %>%
  setView(lng = 0, lat = 30, zoom = 2) %>%
  addPolygons(
    fillColor = ~ pal(Life),
    color = "white",
    fillOpacity = 0.7,
    label = ~labels,
    highlight = highlightOptions(
      color = "black",
      bringToFront = TRUE
    )
  ) %>%
  leaflet::addLegend(
    pal = pal, values = ~Life,
    opacity = 0.7, title = "Life"
  )
```


   

Row {.tabset .tabset-fade}
-------------------------------------

### Definitions

  - **Greenhouse gases** absorb heat (infrared radiation) from the Earth´s surface and then emit it back to the surface, producing the greenhouse effect. In simple words, radiation enters the atmosphere, and gets *traped*. 

  - **GDP** is the gross domestic product, which is the total value of the goods and services produced in a country during a specific period of time. 

  - **Life expectancy** is the average number of years a person can expect to live given certain conditions, i.e. life expectancy by country, sex or other demographic factors.
  


### Average emission of GHG, GDP and Life Expectancy per continent, 2017

```{r}

table<-map@data %>%
  filter(!is.na(GHG)& !is.na(GDP)) %>%
  group_by(continent) %>% 
  summarise(GHG = mean(GHG),
            GDP = mean(GDP),
            "Life Expectancy" = mean(Life))

datatable(table, class = 'cell-border stripe',
          options = list(searching=FALSE, 
                         pageLength=6,
                         lengthChange = FALSE))
```   

### GDP per capita vs Life Expectancy related to GHG emissions per continent, 2017

```{r}

MyPalette = c("#DD8D29", "#E2D200", "#46ACC8", "#E58601", "#B40F20")
map@data %>%
  filter(continent != "Antarctica" & continent != "Seven seas (open ocean)") %>%
ggplot(aes(x=GDP, y=Life),na.rm = TRUE) +
    geom_point(aes(size=GHG, colour= GHG))+
 scale_x_continuous(breaks=c(0,40000,80000,120000),limits=c(-100, 125000))+
  scale_colour_gradientn(limits=c(0, 13000000), breaks=seq(0, max(map$GHG, na.rm = TRUE) + 10, length.out=10), colours = MyPalette)+
  scale_size_continuous(limits=c(0, 13000000), breaks=seq(0, max(map$GHG, na.rm = TRUE) + 10, length.out=10)) +
 guides(color= guide_legend(), size=guide_legend())+
  labs(x='GDP per capita',
      y='Life expectancy')+
  theme(axis.text =element_text(size=10,color="#191919"),
        axis.title=element_text(size=20,color="#191919"),
        panel.background = element_blank(),
        panel.border = element_rect(color="#191919", size=1, fill=NA),
        panel.grid.major =  element_line(colour = "#708090",size=0.1),
        strip.text.x = element_text(size = 10),
        legend.title = element_text(size = 14),
        legend.text = element_text(size = 10))+
theme_light()+
 facet_wrap(~continent, nrow=4)
```

